---
title: "Extra credit"
subtitle: "INFO 2950 - Spring 2023"
author: "Lina Liu"
date: "05/10/23"
format: pdf
editor: visual
---

## Setup

Load packages and data:

```{r}
#| label: load-packages

library(tidyverse)
library(usmap)
library(ggplot2)
library(wesanderson)
library(scales)

#| label: load-data
childcare_costs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-09/childcare_costs.csv')
counties <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-09/counties.csv')
```

## Extra credit

**Research Question:** What is the labor force participation of mothers who have children throughout the United States in the year 2018?

```{r}
#| label: data-wrangling

#| label: labor-county
labor_part_2018 <- childcare_costs |>
  # select relevant columns to use for analysis
  select(county_fips_code, study_year, flfpr_20to64_under6, flfpr_20to64_6to17) |>
  # filter so we only use the year 2018
  filter(study_year == "2018") |>
  # did this justttt in case if there's n/a values
  drop_na() |>
  # combine the column for labor participation of mothers with children under the age
  # of 6 alongside the labor participation of mothers with children 6 to 17.
  pivot_longer(
    cols = -c("county_fips_code", "study_year"),
    names_transform = parse_number,
    values_to = "pct"
  ) |>
  # get rid of the name column, which is kindaaa unnecessary :p
  select(-name) |>
  # because there's two rows of the same area (one for the labor participation
  # percentage of mothers with children under 6 age group and the other one for
  # the 6-17 age group), I decided to average them together to get the total
  # labor participation
  group_by(county_fips_code) |>
  summarize(pct = mean(pct)) |>
  # aligning the fips code of the data so it matches w/ the fips code and column
  # name that the usmap package gave. I added a leading zero in the beginning &
  # changed the column name to fips
  mutate(
    county_fips_code = as.character(county_fips_code),
    fips = if_else(nchar(county_fips_code) == 4,
                               paste0("0", county_fips_code),
                               county_fips_code),
    pct = pct/100)

#| label: labor-states
state_joined <- inner_join(x = childcare_costs, y = counties,
             by = "county_fips_code")

labor_part_2018_states <- state_joined |>
  select(study_year,
         flfpr_20to64_under6,
         flfpr_20to64_6to17,
         state_abbreviation) |>
  filter(study_year == "2018") |>
  drop_na() |>
  pivot_longer(
    cols = -c("study_year", "state_abbreviation"),
    names_transform = parse_number,
    values_to = "pct"
  ) |>
  group_by(state_abbreviation) |>
  summarize(pct = mean(pct)) |>
  mutate(
    state = state_abbreviation,
    pct = pct/100)
```

```{r}
#| label: creating-visualization
#| fig-height: 10
#| fig-width: 12

#| label: visualization-by-county
# I followed this tutorial for the US map:
# https://jtr13.github.io/cc19/different-ways-of-plotting-u-s-map-in-r.html

plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005)  + 
  labs(title = "Labor Force Participation of Mothers throughout the United States
by County",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent) +
# I referred to this source to make my state lines thicker:
# https://stackoverflow.com/questions/72543606/how-to-make-census-region-
# border-thicker-on-usmap-packake-r
  geom_polygon(data = usmapdata::us_map(regions = "states"),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 0.7,
               color = "black")

#| label: visualization-by-state
state_map <- plot_usmap(data = labor_part_2018_states,
           values = "pct",
           linewidth = 0.7,
           labels = TRUE,
           color = "white") +
  labs(title = "Labor Force Participation of Mothers throughout the United States
by States",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent)

# Referred to this as source to make state abbr font smaller:
# https://stackoverflow.com/questions/60806822/how-do-i-change-state-or-counties-
# label-sizes-in-r-with-the-function-usmap
state_map$layers[[2]]$aes_params$size <- 3
print(state_map)
```

**Purpose**

-   The purpose of the visualization is to show how the labor force of the mothers in 2018 compare throughout the United States by different states and counties. Moreover, I thought it was fitting for Mother's day.

-   It helps to answer the question of the labor force participation of mothers who have children throughout the United States in the year 2018.

-   I chose this type of visualization because it incorporates different areas of the United States, which is perfect for the data. It also kinda looked fun and it's something I haven't done before!

Happy Mother's Day!! :)

**Closer view of the map**

The following maps show the labor force participation of mothers by region (for fun!). I only added the code chunk for first map since the code is repetitive.

```{r}
#| label: east-north-central
#| fig-height: 10
#| fig-width: 12

east_north_central <- plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005,
           include = .east_north_central,
           labels = TRUE)  + 
  labs(title = "Labor Force Participation of Mothers in the East North Central region",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent) +
    geom_polygon(data = usmapdata::us_map(regions = "states",
                                          include = .east_north_central),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 1,
               color = "black")

east_north_central$layers[[2]]$aes_params$size <- 8
print(east_north_central)
```

```{r}
#| label: by-region
#| fig-height: 10
#| fig-width: 12
#| echo: false

#| label: east-south-central
east_south_central <- plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005,
           include = .east_south_central,
           labels = TRUE)  + 
  labs(title = "Labor Force Participation of Mothers in the East South Central region",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent)  +
    geom_polygon(data = usmapdata::us_map(regions = "states",
                                          include = .east_south_central),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 1,
               color = "black")

east_south_central$layers[[2]]$aes_params$size <- 8
print(east_south_central)

#| label: mid-west-region
midwest_region <- plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005,
           include = .midwest_region,
           labels = TRUE)  + 
  labs(title = "Labor Force Participation of Mothers in the Mid-west region",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent) +
    geom_polygon(data = usmapdata::us_map(regions = "states",
                                          include = .midwest_region),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 1,
               color = "black")

midwest_region$layers[[2]]$aes_params$size <- 8
print(midwest_region)

#| label: mid-atlantic
midatlantic <- plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005,
           include = .mid_atlantic,
           labels = TRUE)  + 
  labs(title = "Labor Force Participation of Mothers in the Mid-Atlantic region",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent) +
    geom_polygon(data = usmapdata::us_map(regions = "states",
                                          include = .mid_atlantic),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 1,
               color = "black")

midatlantic$layers[[2]]$aes_params$size <- 8
print(midatlantic)

#| label: mountain
mountain <- plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005,
           include = .mountain,
           labels = TRUE)  + 
  labs(title = "Labor Force Participation of Mothers in the Mountain region",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent) + 
  geom_polygon(data = usmapdata::us_map(regions = "states",
                                          include = .mountain),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 1,
               color = "black")

mountain$layers[[2]]$aes_params$size <- 8
print(mountain)

#| label: new-england
new_england <- plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005,
           include = .new_england,
           labels = TRUE)  + 
  labs(title = "Labor Force Participation of Mothers in the New England region",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent) +
    geom_polygon(data = usmapdata::us_map(regions = "states",
                                          include = .new_england),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 1,
               color = "black")

new_england$layers[[2]]$aes_params$size <- 8
print(new_england)

#| label: north-east-region
north_east_region <- plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005,
           include = .northeast_region,
           labels = TRUE)  + 
  labs(title = "Labor Force Participation of Mothers in the Northeast region",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent)  +
    geom_polygon(data = usmapdata::us_map(regions = "states",
                                          include = .northeast_region),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 1,
               color = "black")

north_east_region$layers[[2]]$aes_params$size <- 8
print(north_east_region)

#| label: north-central-region
north_central_region <- plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005,
           include = .north_central_region,
           labels = TRUE)  + 
  labs(title = "Labor Force Participation of Mothers in the North Central region",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent) +
    geom_polygon(data = usmapdata::us_map(regions = "states",
                                          include = .north_central_region),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 1,
               color = "black")

north_central_region$layers[[2]]$aes_params$size <- 8
print(north_central_region)

#| label: pacific
pacific <- plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005,
           include = .pacific,
           labels = TRUE)  + 
  labs(title = "Labor Force Participation of Mothers in the Pacific region",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent) + 
  geom_polygon(data = usmapdata::us_map(regions = "states",
                                          include = .pacific),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 1,
               color = "black")

pacific$layers[[2]]$aes_params$size <- 8
print(pacific)

#| label: south-atlantic
south_atlantic <- plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005,
           include = .south_atlantic,
           labels = TRUE)  + 
  labs(title = "Labor Force Participation of Mothers in the South Atlantic region",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent) +
    geom_polygon(data = usmapdata::us_map(regions = "states",
                                          include = .south_atlantic),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 1,
               color = "black")

south_atlantic$layers[[2]]$aes_params$size <- 6.5
print(south_atlantic)

#| label: south-region
south_region <- plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005,
           include = .south_region,
           labels = TRUE)  + 
  labs(title = "Labor Force Participation of Mothers in South region",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent) +
    geom_polygon(data = usmapdata::us_map(regions = "states",
                                          include = .south_region),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 1,
               color = "black")

south_region$layers[[2]]$aes_params$size <- 8
print(south_region)

#| label: west-north-central
west_north_central <- plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005,
           include = .west_north_central,
           labels = TRUE)  + 
  labs(title = "Labor Force Participation of Mothers West North Central region",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent) +
    geom_polygon(data = usmapdata::us_map(regions = "states",
                                          include = .west_north_central),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 1,
               color = "black")

west_north_central$layers[[2]]$aes_params$size <- 8
print(west_north_central)

#| label: west-region
west_region <- plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005,
           include = .west_region,
           labels = TRUE)  + 
  labs(title = "Labor Force Participation of Mothers in the West region",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent) +
    geom_polygon(data = usmapdata::us_map(regions = "states",
                                          include = .west_region),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 1,
               color = "black")

west_region$layers[[2]]$aes_params$size <- 8
print(west_region)

#| label: west-south-central
west_south_central <- plot_usmap(data = labor_part_2018,
           values = "pct",
           linewidth = 0.0005,
           include = .west_south_central,
           labels = TRUE)  + 
  labs(title = "Labor Force Participation of Mothers in the West South Central region",
       subtitle = "For the year 2018",
       caption = "Source: National Database of Childcare Prices",
       fill = "Labor Force Participation") + 
  theme(panel.background = element_blank(),
        legend.position = "left") +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100,
                                             type = "continuous"),
                       labels = percent)  +
    geom_polygon(data = usmapdata::us_map(regions = "states",
                                          include = .west_south_central),
               aes(x, y, group = group),
               fill = NA,
               linewidth = 1,
               color = "black")

west_south_central$layers[[2]]$aes_params$size <- 8
print(west_south_central)
```
